Fixed-Parameter Tractable Sampling for RNA Design with Multiple Target Structures

### 2018-02-28 RNARedPrint-Bioinfo2018.tex
Typos

%%% macros for notation in DP framework
\newcommand{\network}{\mathcal{N}}
\newcommand{\val}{\bar S} % valuation aka assignment
\newcommand{\dep}{\operatorname{dep}}
\newcommand{\energy}[1]{\operatorname{e}_{#1}}
\newcommand{\numberof}{\operatorname{\#}}
\newcommand{\partfun}[1]{\mathcal{Z}_{#1}}
\newcommand{\separator}[2]{\operatorname{sep}(#1,#2)}
\newcommand{\difference}[2]{\operatorname{diff}(#1 \rightarrow #2)}
\newcommand{\real}{\mathbb{R}}
\newcommand{\genmarg}[1]{(\!|\!#1\!|\!)}
\newcommand{\gencomb}[1]{\langle\!|#1|\!\rangle}
\newcommand{\Message}[2]{m_{#1\rightarrow #2}}

@@

%
The \Def{energy $\energy{\network}(S)$ of a sequence} $S$ in a network
$\network$ is defined as sum of the values of all functions in
$f\in\F$ evaluated for $S$, i.e.
$\energy{\network}(S) := \sum_{f\in F} \evalfor{f}{S}.$

The network energy $\energy{\network}(S)$ corresponds to the energy in
Eq.~$(\ref{eq:mainproblem}),$ where this energy is modeled as sum of
the functions in $\F$. Consequently, $\partfun{\R}$ of
Eq.~$(\ref{eq:mainproblem})$ is modeled as network partition function
$\partfun{\network} := \sum_{S}\exp(-\beta\energy{\network}(S)) = \sum_{S}\prod_{f\in F} \exp( -\beta\cdot
\evalfor{f}{S} ).$

@@

\begin{algorithm}[t]
  \KwData{Cluster tree $(T,\chi,\phi)$} \KwResult{Messages
    $\Message{u}{v}$ for all $(u\to{}v)\in T$; i.e.~partition
    functions of the subtrees of all $v$ for all possible partial
    sequences determining exactly the positions $\separator{u}{v}$.}
 \For{$u\to{}v\in T$ in postorder}{
  \For{$\val\in\partseqs(\separator{u}{v})$}{
    $\evalfor{\Message{u}{v}}{\val} \gets 0$\;
    \For{$\val'\in\partseqs(\difference{u}{v})$}{
     $p \gets$ product( $exp(-\beta \evalfor{f}{\substitute{\val}{\val'}})$ for $f\in \phi(u)$ )\;
     $p \gets p\ \cdot\ $product( $\evalfor{\Message{w}{u}}{\substitute{\val}{\val'}}$ for $(w\to{}u)\in T$ )\;
     $\evalfor{\Message{u}{v}}{\val} \gets \evalfor{\Message{u}{v}}{\val} + p$\;
   }
  }
  \Return {$m$}\;
  }
 \caption{FPT computation of the partition function using
   dynamic programming (CTE). }\label{alg:pf}
\end{algorithm}

@@

     $p \gets$ product( $exp(-\beta \evalfor{f}{\substitute{\val}{\val'}})$ for $f\in \phi(u)$ )\;
     $p \gets p\ \cdot\ $product( $\evalfor{\Message{w}{u}}{\substitute{\val}{\val'}}$ for $(w\to{}u)\in T$ )\;
%
%     $p \gets 1$\;
%     \For{$f\in \phi(v)$}{
%  	      $p \gets p \cdot \exp(-\beta \evalfor{f}{\substitute{\val}{\val'}})$\;
%     }
%     \For{$v\to{}w\in T$}{
%  	      $p \gets p \cdot \evalfor{\Message{v}{w}}{\substitute{\val}{\val'}}$\;
%     }
  	  $r \gets r - p$\;
  	  \If{$r<0$}{
  	    $a_{\rm res} \gets \substitute{\val}{\val'}$\;
  	  \For{$(v\to{}w)\in T$}{
  	      $a_{\rm res} \gets \substitute{a_{\rm res}}{\Sample(v,\substitute{\val}{\val'};T,\chi,\phi,m)}$\;
     }
  	  \Return {$a_{\rm res}$}
  	  \;}
 }
 }
 \caption{Stochastic backtrack algorithm for partial sequences in the Boltzmann distribution.}\label{alg:sampling}
\end{algorithm}
%For example, for design to a single non-crossing target in the nearest neighbor energy model, the complexity of Algs.~\ref{alg:pf} and \ref{alg:sampling} degrades to linear time (as reported for IncaRNAtion), since there are only $O(n)$ many functions and nodes (in the sequence length $n$). Moreover, the dependencies are tree-like owing to the tree-like non-crossing structure; this implies a constantly bounded maximum treewidth.

>>>>

%%% macros for notation in DP framework
\newcommand{\network}{\mathcal{N}}
\newcommand{\val}{\bar S} % valuation aka assignment
\newcommand{\dep}{\operatorname{dep}}
\newcommand{\energy}[1]{\operatorname{e}_{#1}}
\newcommand{\numberof}{\operatorname{\#}}
\newcommand{\partfun}[1]{Z_{#1}}
\newcommand{\separator}[2]{\operatorname{sep}(#1,#2)}
\newcommand{\difference}[2]{\operatorname{diff}(#1 \rightarrow #2)}
\newcommand{\real}{\mathbb{R}}
\newcommand{\genmarg}[1]{(\!|\!#1\!|\!)}
\newcommand{\gencomb}[1]{\langle\!|#1|\!\rangle}
\newcommand{\Message}[2]{m_{#1\rightarrow #2}}

@@

%
The \Def{energy $\energy{\network}(S)$ of a sequence} $S$ in a network
$\network$ is defined as sum of the values of all functions in
$f\in\F$ evaluated for $S$, i.e.
$\energy{\network}(S) := \sum_{f\in\F} \evalfor{f}{S}.$

The network energy $\energy{\network}(S)$ corresponds to the energy in
Eq.~$(\ref{eq:mainproblem}),$ where this energy is modeled as sum of
the functions in $\F$. Consequently, $\partfun{\R}$ of
Eq.~$(\ref{eq:mainproblem})$ is modeled as network partition function
$\partfun{\network} := \sum_{S}\exp(-\beta\energy{\network}(S)) = \sum_{S}\prod_{f\in\F} \exp( -\beta\cdot
\evalfor{f}{S} ).$

@@

\begin{algorithm}[t]
  \KwData{Cluster tree $(T,\chi,\phi)$} \KwResult{Messages
    $\Message{u}{v}$ for all $(u\to{}v)\in T$; i.e.~partition
    functions of the subtrees of all $v$ for all possible partial
    sequences determining exactly the positions $\separator{u}{v}$.}
 \For{$u\to{}v\in T$ in postorder}{
  \For{$\val\in\partseqs(\separator{u}{v})$}{
    $x\gets 0$\;
    \For{$\val'\in\partseqs(\difference{u}{v})$}{
     $p \gets$ product( $exp(-\beta \evalfor{f}{\substitute{\val}{\val'}})$ for $f\in \phi(u)$ )\\
     ${}\quad\qquad \cdot\ $product( $\evalfor{\Message{w}{u}}{\substitute{\val}{\val'}}$ for $(w\to{}u)\in T$ )\;
     $x \gets x + p$\;
   }
   $\evalfor{\Message{u}{v}}{\val} \gets x$\;
  }
  \Return {$m$}\;
  }
 \caption{FPT computation of the partition function using
   dynamic programming (CTE). }\label{alg:pf}
\end{algorithm}

@@

     $p \gets$ product( $exp(-\beta \evalfor{f}{\substitute{\val}{\val'}})$ for $f\in \phi(u)$\\
     ${}\quad\qquad \cdot\ $product( $\evalfor{\Message{w}{u}}{\substitute{\val}{\val'}}$ for $(w\to{}u)\in T$ )\;
%
%     $p \gets 1$\;
%     \For{$f\in \phi(v)$}{
%  	      $p \gets p \cdot \exp(-\beta \evalfor{f}{\substitute{\val}{\val'}})$\;
%     }
%     \For{$v\to{}w\in T$}{
%  	      $p \gets p \cdot \evalfor{\Message{v}{w}}{\substitute{\val}{\val'}}$\;
%     }
  	  $r \gets r - p$\;
  	  \If{$r<0$}{
  	    $\val_{\rm res} \gets \substitute{\val}{\val'}$\;
  	  \For{$(v\to{}w)\in T$}{
  	      $\val_{\rm res} \gets \substitute{\val_{\rm res}}{\Sample(v,\substitute{\val}{\val'};T,\chi,\phi,m)}$\;
     }
  	  \Return {$\val_{\rm res}$}
  	  \;}
 }
 }
 \caption{Stochastic backtrack algorithm for partial sequences in the Boltzmann distribution.}\label{alg:sampling}
\end{algorithm}
%For example, for design to a single non-crossing target in the nearest neighbor energy model, the complexity of Algs.~\ref{alg:pf} and \ref{alg:sampling} degrades to linear time (as reported for IncaRNAtion), since there are only $O(n)$ many functions and nodes (in the sequence length $n$). Moreover, the dependencies are tree-like owing to the tree-like non-crossing structure; this implies a constantly bounded maximum treewidth.
